---
title: "Vignette for reading and summarizing data from API's"
author: "Demetrios Samaras"
date: "2023-06-16"
output:
   html_document:
    toc: TRUE
    toc_depth: 3
    toc_float: FALSE
    code_folding: hide
    theme: readable
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE)

##api key = DrI5M1TDWji8nJS9aZDfUAdR_6iqhFzm
```

# Purpose 

##### The purpose of this Vignette is to summarize how to interact with, pull data and then analyze the data from the polygon.io finacial data API [https://polygon.io](https://polygon.io)

# Required packages 

##### The following packages will be required 

  * **tidyverse**
    + We will use a number of the functions from the tidyverse
  * **jsonlite**
    + This will help up interact with the API
  * **httr**
    + for the `GET` function 
  * **ggplot2**
    + This will be required to create the data visualizations
    
# Steps for querying endpoints  

#### Step 0 

Make sure all of the required packages are installed using the `install.packages()` function and library them using `library()`

```{r library functions, echo=TRUE, eval=TRUE}

library(tidyverse)
library(jsonlite)
library(httr)
library(ggplot2)

```

#### Step 1 Contact the API 

```{r contact , echo = TRUE, eval = FALSE}
## this is a function to contact the API based on the ticker of interest
ticker <- function(ticker){
  
  link
  
 response <-GET("https://api.polygon.io/v2/aggs/ticker/BA/range/1/day/2023-01-09/2023-01-09?adjusted=true&sort=asc&limit=120&apiKey=DrI5M1TDWji8nJS9aZDfUAdR_6iqhFzm")

response
}
  
```




 













